Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-08 prop_letter.tex
prop_letter.tex Simplify the TeX code...
prop_letter.tex \section{Born} Updating...

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{\vphantom{|}}\middle|\mathrel{\vphantom{|}}{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\mathrm{L{#1}}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\mathrm{R{#1}}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{\mathrm{d}}
\newcommand{\rmi}{\mathrm{i}}
\lineskiplimit=1pt
\newcommand{\iotal}[1][]{\ensuremath{\iota^{\mathrm{L{#1}}}}}
\newcommand{\iotar}[1][]{\ensuremath{\iota^{\mathrm{R{#1}}}}}

@@

\yutsung{Notice that the use of quantum is minumalized in this section.
In another word, the results of this section should apply to both
quantum and classical case.}

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates of the quantum probability measures first, and then introduce
the Born rule which induces the quantum probability measures for each
state. When we shift our attention to QIVPMs, we also want to construct
a QIVPMs according to a simple rule. Since the ranges of the quantum
probability measures and QIVPMs are $\left[0,1\right]$ and $\mathscr{I}$,
respectively, if there is a function~$\iota$ mapping a number~$x\in\left[0,1\right]$
to $\iota\left(x\right)\in\mathscr{I}$, then we can concatenate it
with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
and get a QIVPM~$\bar{\mu}:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$.
Although their concatenation $\iota\circ\mu$ is always a function
from $\events$ to $\mathscr{I}$, to guarantee $\iota\circ\mu$ is
a QIVPM, $\iota$ needs to satisfy some extra conditions.

\begin{thm}If $\iota\circ\mu:\events\rightarrow\mathscr{I}$ is always
a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
in any Hilbert space $\Hilb$, then $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
must subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations} and satisfying for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\subseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{thm}

If $\iota$ fails to satisfy the first three constraints, Eqs.~(\ref{eq:iota-constraints}),
it is easy to see that the concatenated function~$\iota\circ\mu$
may not be a QIVPM. Suppose $\iota$ fails to satisfy Eq.~(\ref{eq:iota-Inclusion}),
i.e., there exists three numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but 
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion-not}
\end{equation}
Consider the state
\begin{equation}
\begin{aligned}\rho=\vphantom{x_{0}} & x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}\\
 & +\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\textrm{ .}
\end{aligned}
\end{equation}
Let $P_{0}=\proj{0}+\proj{2}$, $P_{1}=\proj{1}+\proj{2}$. Then,
the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})

\begin{eqnarray}
 &  & \begin{aligned} & \mathrel{\hphantom{=}}\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\end{aligned}
\nonumber \\
 &  & \begin{aligned} & =\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\\
 & \nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)
\end{aligned}
\\
 &  & \begin{aligned} & =\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\textrm{ ,}\end{aligned}
\nonumber 
\end{eqnarray}
and cannot be a QIVPM. Therefore, $\iota$ must satisfy Eqs.~(\ref{eq:iota-constraints})
and Eq.~(\ref{eq:iota-Inclusion}) to guarantee $\iota\circ\mu$
is always a QIVPM.

>>>>

\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{framed}
\newcommand{\amr}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{green}{Amr says: #1}\end{minipage}\end{framed}}
\newcommand{\yutsung}[1]{\begin{framed}\begin{minipage}{0.9\linewidth}\color{purple}{Yu-Tsung says: #1}\end{minipage}\end{framed}}
\newcommand{\set}[2]{\ensuremath{\left\{ {#1}\mathrel{}\middle|\mathrel{}{#2}\right\} }}
\newcommand{\Tr}{\ensuremath{\mathop{\mathrm{Tr}}\nolimits}}
\allowdisplaybreaks
\newcommand{\coreBorn}{\ensuremath{\overline{\Hilb}}}
\newcommand{\mul}[1][]{\ensuremath{\mu^{\mathrm{L{#1}}}}}
\newcommand{\mur}[1][]{\ensuremath{\mu^{\mathrm{R{#1}}}}}
\setcounter{secnumdepth}{3}
% https://tex.stackexchange.com/questions/345284/can-i-have-an-almost-non-breaking-space-in-latex
% Change non-breaking space to \nolinebreak[1] 
% Hence, it will not hyphenate too much surrounding words.
\newcommand{\rmd}{\mathrm{d}}
\newcommand{\rmi}{\mathrm{i}}
\lineskiplimit=1pt
\newcommand{\iotal}[1][]{\ensuremath{\iota^{\mathrm{L{#1}}}}}
\newcommand{\iotar}[1][]{\ensuremath{\iota^{\mathrm{R{#1}}}}}

@@

\yutsung{Notice that the use of quantum is minumalized in this section.
In another word, the results of this section should apply to both
quantum and classical IVPMs.}

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates of the quantum probability measures first, and then introduce
the Born rule which induces the quantum probability measures for each
state. When we shift our attention to QIVPMs, we also want to construct
a QIVPMs according to a simple rule. Since the ranges of the quantum
probability measures and QIVPMs are $\left[0,1\right]$ and $\mathscr{I}$,
respectively, if there is a function~$\iota$ mapping a number~$x\in\left[0,1\right]$
to $\iota\left(x\right)\in\mathscr{I}$, then we can concatenate it
with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
and get a function~$\bar{\mu}:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$.
Although $\bar{\mu}$ has the same the domain and range, $\bar{\mu}$
may or may not be a QIVPM. For example, the following theorem highlight
the situation when $\bar{\mu}$ cannot be a QIVPM.

\begin{thm}If there exists three numbers~$x_{0}$, $x_{1}$, and
$x_{2}\in\left[0,1\right]$ such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$,
but 
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion-not}
\end{equation}
Then, there is a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
such that $\iota\circ\mu$ is not a QIVPM, where $\circ$ is the function
concatenation.\end{thm}

Consider the state
\begin{equation}
\begin{aligned}\rho={} & x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}\\
 & +\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\textrm{ .}
\end{aligned}
\end{equation}
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray}
 &  & \begin{aligned} & \mathrel{\hphantom{=}}\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\end{aligned}
\nonumber \\
 &  & \begin{aligned} & =\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\\
 & \nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)
\end{aligned}
\\
 &  & \begin{aligned} & =\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\textrm{ ,}\end{aligned}
\nonumber 
\end{eqnarray}
and cannot be a QIVPM.

In other words, even if $\iota\circ\mu$ has the same domain and range
as a QIVPM, we may need to verify whether $\iota\circ\mu$ is a QIVPM
individually. For example, consider the following easy example. In
the four-dimensional Hilbert space, 
\begin{equation}
\iota\left(x\right)=\begin{cases}
\imposs & \textrm{, if }x=0\textrm{ ;}\\
\left[0,\tfrac{1}{2}\right] & \textrm{, if }x\in\left(0,\tfrac{1}{4}\right]\textrm{ ;}\\
\unknown & \textrm{, if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\textrm{ ;}\\
\left[\tfrac{1}{2},1\right] & \textrm{, if }x\in\left[\tfrac{3}{4},1\right)\textrm{ ;}\\
\necess & \textrm{, if }x=1\textrm{ .}
\end{cases}
\end{equation}
and $\rho=\left(\proj{0}+\proj{1}+\proj{2}+\proj{3}\right)/4$ induce
a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}$ which is a
non-trivial fact~\citep{HOSTunpublished}. In order to verify whether
$\iota\circ\mu$ is a QIVPM easier, we want to impose some conditions
on $\iota$. Notice that our counterexamples are in a four-dimensional
Hilbert space, the conditions for $\iota$ for a Hilbert space of
dimension~$d\le3$ is not as strict as the general conditions.

\begin{thm}In a Hilbert space of dimension~$d\le3$, if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfying for any three numbers~$x_{0}$
and $x_{1}\in\left[0,1\right]$, $x_{0}+x_{1}\in\left[0,1\right]$
implies
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\textrm{ .}\label{eq:iota-add}
\end{equation}
Then, $\iota\circ\mu:\events\rightarrow\mathscr{I}$ is always a QIVPM
for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$.\end{thm}

Although their concatenation $\iota\circ\mu$ is always a function
from $\events$ to $\mathscr{I}$, to guarantee $\iota\circ\mu$ is
a QIVPM, $\iota$ needs to satisfy some extra conditions.

\begin{thm}If $\iota\circ\mu:\events\rightarrow\mathscr{I}$ is always
a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
in any Hilbert space $\Hilb$, then $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
must subject to the following constraints:\begin{subequations}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations} and satisfying for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\subseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{thm}
